---
title: Tidy Tuesday January 8, 2019
---

```{r}
knitr::opts_chunk$set(
  echo = TRUE
)

library(tidyverse)
```

Doing this in favorite Gather -> Cook -> Analyze -> Present format, as implemented in my [represtools package](https://cran.r-project.org/package=represtools). We'll note be presenting results - this is just EDA for fun.

## Gather

```{r}
url <- 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-01-08/IMDb_Economist_tv_ratings.csv'

tbl_raw <- read_csv(url)
```

Raw data in hand, I start to cook what I have. This is largely normalization. It's overkill, but I'm fussy and struggle to look at data any other way. 

## Cook: Title table

The raw data is a mashup of title, season and genre. We can split those out.

Title is easy to pick up. Simply `select()` the ID and the title and then run `unique()`.

```{r}
tbl_title <- tbl_raw %>% 
  select(titleId, title) %>% 
  unique()
```

## Cook: Check genre change

We might expect that genre is a property of the title. That is, each title has one genre value which remains the same over time. That's not actually the case.

```{r}
tbl_genre_switch <- tbl_raw %>% 
  select(titleId, genres) %>% 
  group_by(titleId) %>% 
  summarise(num_genre = n_distinct(genres)) %>% 
  ungroup() %>% 
  filter(num_genre > 1) %>% 
  inner_join(tbl_raw, by = 'titleId') %>% 
  select(title, date, genres)

tbl_genre_switch
```

Twin Peaks has different genre information between its two incarnations. Seems it morphed from a mystery to fantasy. Not a big deal and one that we could manually correct. However, I'm not going to touch it and will simply 

## Cook: Seasons

This is simply dropping the genre and title fields. In addition, we will add a column for the year, which will help us visualize later.

```{r}
tbl_season <- tbl_raw %>% 
  select(-genres, -title) %>% 
  mutate(
      year = lubridate::year(date)
    , semester = lubridate::semester(date, with_year = TRUE)
  )
```

## Cook: Split genres

The genres are contained in a vector, which is not typically how I deal with normalized data. We can use `separate` to change that single column into multiple, but first we need to know how many columns to create.  `str_count` of the comma separator will do that. We need to make sure to add 1 to the result.

```{r}
tbl_genre <- tbl_raw %>% 
  select(titleId, genre = genres)

max_genres <- tbl_genre %>% 
  mutate(
    commas = str_count(genre, ',')
  ) %>% 
  summarise(
    max_col = max(commas) + 1
  ) %>% 
  unlist()
```

Now, we can split the genre column into 3. We will be warned that not every row has all three. Cool. We'll filter them out, after using gather to translate our data into a table which stores a titleId and a genre value.

```{r }
tbl_genre <- tbl_genre %>% 
  tidyr::separate(genre, into = as.character(seq_len(max_genres)), sep = ',') %>% 
  gather(id, genre, -titleId) %>% 
  filter(!is.na(genre)) %>% 
  select(-id) %>% 
  unique()
```

In structuring the data this way, we are assuming that we can map the ratings data to genre, irrespective of season. That's not 100% accurate, as we saw earlier. The practical effect here means that Twin Peaks will map to four different genres

Cook: Consolidate genres

There are many genres, so let's focus on the top 8. `forcats` will do this. We add this as a new column so that we can see the more granular value for genre, if we'd like.

```{r}
tbl_genre <- tbl_genre %>% 
  mutate(
    genre_other = forcats::fct_lump(genre, n = 8)
  )
```

## Analyze: 1 - Genre over time

```{r}
tbl_genre_time <- tbl_genre %>% 
  inner_join(tbl_season %>% select(titleId, date, year, av_rating), by = 'titleId')
```

```{r }
tbl_genre_time %>% 
  group_by(genre_other, year) %>% 
  summarise(
    count = n()
  ) %>% 
  ggplot(aes(year, count, color = genre_other)) + 
  geom_point()
```

```{r}
tbl_genre_time %>% 
  ggplot(aes(year, av_rating, color = genre_other)) + 
  geom_point() +
  geom_smooth() +
  facet_wrap(~genre_other)
```

```{r}
tbl_genre_time %>% 
  mutate(year = as.factor(year)) %>% 
  ggplot(aes(year, av_rating, color = genre_other)) + 
  geom_boxplot() +
  facet_wrap(~genre_other)
```

```{r}
tbl_genre_single <- tbl_genre %>% 
  group_by(titleId) %>% 
  summarise(
    num_genres = n()
  ) %>% 
  filter(
    num_genres == 1
  ) %>% 
  inner_join(tbl_genre)

tbl_pure_drama <- tbl_genre_single %>% 
  filter(genre == "Drama") %>% 
  select(titleId) %>% 
  inner_join(tbl_season)
```

```{r}
tbl_pure_drama %>% 
  mutate(year = as.factor(year)) %>% 
  ggplot(aes(year, av_rating)) + 
  geom_boxplot()
```

```{r}
tbl_pure_drama %>% 
  ggplot(aes(year, av_rating)) + 
  geom_point() + 
  geom_smooth()
```

```{r}
tbl_pure_drama %>% 
  ggplot(aes(year)) + 
  geom_bar()
```

Non-comedy drama

```{r}
tbl_drama_non_comedy <- tbl_genre %>% 
  anti_join(
      tbl_genre %>% filter(genre == 'Comedy')
    , by = 'titleId'
  ) %>% 
  select(titleId) %>% 
  unique() %>% 
  inner_join(tbl_season)
```

```{r}
tbl_drama_non_comedy %>% 
  ggplot(aes(year)) + 
  geom_bar()
```

```{r}
tbl_drama_non_comedy %>% 
  mutate(year = as.factor(year)) %>% 
  ggplot(aes(year, av_rating)) + 
  geom_boxplot()

tbl_drama_non_comedy %>% 
  ggplot(aes(year, av_rating)) + 
  geom_point() + 
  geom_smooth()
```

It looks as though in recent years,the overall quality has improved, but there are more negative outliers. That is, dramas are marginally better overall, but the bad ones are worse. We can get a sense for this by examining the skewness of the samples. It's clear that every year is negatively skewed

```{r}
library(e1071)

tbl_drama_non_comedy_year <- tbl_drama_non_comedy %>% 
  group_by(year) %>% 
  summarise_at(
      'av_rating'
    , .funs = funs(mean, median, sd, var, skewness, kurtosis)
  )

tbl_drama_non_comedy_year %>% 
  ggplot(aes(year, skewness)) +
  geom_point() +
  geom_smooth()
```

