---
title: Tidy Tuesday January 8, 2019
---

```{r}
knitr::opts_chunk$set(
  echo = TRUE
)

library(tidyverse)
```

Doing this in favorite Gather -> Cook -> Analyze -> Present format, as implemented in my [represtools package](https://cran.r-project.org/package=represtools). We'll note be presenting results - this is just EDA for fun.

## Gather

```{r}
url <- 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-01-08/IMDb_Economist_tv_ratings.csv'

tbl_raw <- read_csv(url)
```

Raw data in hand, I start to cook what I have. This is largely normalization. It's overkill, but I'm fussy and struggle to look at data any other way. 

## Cook: Title table

The raw data is a mashup of title, season and genre. We can split those out.

Title is easy to pick up. Simply `select()` the ID and the title and then run `unique()`.

```{r}
tbl_title <- tbl_raw %>% 
  select(titleId, title) %>% 
  unique()
```

## Cook: Check genre change

We might expect that genre is a property of the title. That is, each title has one genre value which remains the same over time. That's not actually the case.

```{r}
tbl_genre_switch <- tbl_raw %>% 
  select(titleId, genres) %>% 
  group_by(titleId) %>% 
  summarise(num_genre = n_distinct(genres)) %>% 
  ungroup() %>% 
  filter(num_genre > 1) %>% 
  inner_join(tbl_raw, by = 'titleId') %>% 
  select(title, date, genres)

tbl_genre_switch
```

Twin Peaks has different genre information between its two incarnations. Seems it morphed from a mystery to fantasy. Not a big deal and one that we could manually correct. However, I'm not going to touch it and will simply 

## Cook: Seasons

This is simply dropping the genre and title fields. In addition, we will add a column for the year, which will help us visualize later.

```{r}
tbl_season <- tbl_raw %>% 
  select(-genres, -title) %>% 
  mutate(
      year = lubridate::year(date)
    , semester = lubridate::semester(date, with_year = TRUE)
  )
```

## Cook: Split genres

The genres are contained in a vector, which is not typically how I deal with normalized data. We've seen earlier that at least one title switches genre

```{r}
tbl_genre <- tbl_raw %>% 
  select(titleId, genre = genres)

max_genres <- tbl_genre %>% 
  mutate(
    commas = str_count(genre, ',')
  ) %>% 
  summarise(
    max_col = max(commas) + 1
  ) %>% 
  unlist()
```

```{r }
tbl_genre <- tbl_genre %>% 
  tidyr::separate(genre, into = as.character(seq_len(max_genres)), sep = ',') %>% 
  gather(id, genre, -titleId) %>% 
  filter(!is.na(genre)) %>% 
  select(-id) %>% 
  unique()
```

Cook: Consolidate genres

```{r}
tbl_genre <- tbl_genre %>% 
  mutate(
    genre_other = forcats::fct_lump(genre, n = 8)
  )
```

## Analyze: 1 - Genre over time

```{r}
tbl_genre_time <- tbl_genre %>% 
  inner_join(tbl_rating %>% select(titleId, date, year, av_rating), by = 'titleId')
```

```{r }
tbl_genre_time %>% 
  group_by(genre_other, year) %>% 
  summarise(
    count = n()
  ) %>% 
  ggplot(aes(year, count, color = genre_other)) + 
  geom_point()
```

```{r}
tbl_genre_time %>% 
  ggplot(aes(year, av_rating, color = genre_other)) + 
  geom_point() +
  geom_smooth() +
  facet_wrap(~genre_other)
```

```{r}
tbl_genre_time %>% 
  mutate(year = as.factor(year)) %>% 
  ggplot(aes(year, av_rating, color = genre_other)) + 
  geom_boxplot() +
  facet_wrap(~genre_other)
```

```{r}
tbl_genre_single <- tbl_genre %>% 
  group_by(titleId) %>% 
  summarise(
    num_genres = n()
  ) %>% 
  filter(
    num_genres == 1
  ) %>% 
  inner_join(tbl_genre)

tbl_pure_drama <- tbl_genre_single %>% 
  filter(genre == "Drama") %>% 
  select(titleId) %>% 
  inner_join(tbl_rating)
```

```{r}
tbl_pure_drama %>% 
  mutate(year = as.factor(year)) %>% 
  ggplot(aes(year, av_rating)) + 
  geom_boxplot()
```

```{r}
tbl_pure_drama %>% 
  ggplot(aes(year, av_rating)) + 
  geom_point() + 
  geom_smooth()
```

```{r}
tbl_pure_drama %>% 
  ggplot(aes(year)) + 
  geom_bar()
```

Non-comedy drama

```{r}
tbl_drama_non_comedy <- tbl_genre %>% 
  anti_join(
      tbl_genre %>% filter(genre == 'Comedy')
    , by = 'titleId'
  ) %>% 
  select(titleId) %>% 
  unique() %>% 
  inner_join(tbl_rating)
```

```{r}
tbl_drama_non_comedy %>% 
  ggplot(aes(year)) + 
  geom_bar()
```

```{r}
tbl_drama_non_comedy %>% 
  mutate(year = as.factor(year)) %>% 
  ggplot(aes(year, av_rating)) + 
  geom_boxplot()

tbl_drama_non_comedy %>% 
  ggplot(aes(year, av_rating)) + 
  geom_point() + 
  geom_smooth()
```

